Using Scipion for Stream Image Processing at Cryo-EM Facilities

### 2018-09-24 Review/4.future.tex
to home

\subsection{Future Developments}

Currently, \scipion includes the appropriate code (python wrappers) to talk to the integrated EM packages (Xmipp, Eman, Relion, ...) at the time of the release. Therefore, an update of any of the integrated EM packages immediately after the \scipion release will not be available until next \scipion release (we are aiming for one major release per year). To decouple EM-package releases from \scipion releases, we are working on making the wrappers for EM packages  independent from \scipion. Plans for ensuring this independence include: (1) wrapper installation through pip (\url{https://pip.pypa.io/en/stable/}), (2) reimplementation of the wrappers as plug-ins that will be automatically detected and added to the application menu and (3) develop of standalone IO  and visualization libraries that make \scipion fully independent from other EM packages.

@@

>>>>

\section{Future Developments and Usage}

\subsection{Future Developments}
Currently, \scipion includes the appropriate code (python wrappers) to talk to the integrated EM packages (Xmipp, Eman, Relion, ...) at the time of the release. Therefore, an update of any of the integrated EM packages immediately after the \scipion release will not be available until next \scipion release (we are aiming for one major release per year). To decouple EM-package releases from \scipion releases, we are working on making the wrappers for EM packages  independent from \scipion. Plans for ensuring this independence include: (1) wrapper installation through pip (\url{https://pip.pypa.io/en/stable/}), (2) reimplementation of the wrappers as plug-ins that will be automatically detected and added to the application menu and (3) develop of standalone IO  and visualization libraries that make \scipion fully independent from other EM packages.

@@

\subsection{Usage}
In the context of free and open source projects it is difficult to create usage statistics that are accurate. But certainly, if we want \scipion to succeed, we need to 
know how frequent the different protocols are used. This is the reason why we have recently developed our \textit{Scipion Usage Data Collector} that monitors -if activated by the user- the usage of the different protocols and send -via HTTP protocol- the information to the developer's team. This information is anonymous and cannot be used to identify the original computer in which \scipion has been executed.

Using the HTTP header it is possible to guess the country in which \scipion has been installed. With these information we have created the usage map and table available at URL \url{http://scipion.i2pc.es/report_protocols/scipionUsage/} and  \url{http://scipion.i2pc.es/report_protocols/protocolTable/}. The data corresponds on ``global'' \scipion usage rather than ``facility'' usage but there 
is a one to one relationship between the  countries with highest number of \scipion projects and the countries with \scipion based EM facilities. The only exception to the rule is Canada where two of the original \scipion developers are setting up a new facility (McGill University).

Note that, in a effort to report real use and skip test projects, the number of projects in the map refers to the number of non empty \scipion projects updated more than once. That is, 
\scipion projects that have been opened in at least two different days. Another constraint is that developers computers are filtered out.

### 2018-09-24 Review/5.conclusions.tex
to home

 Workflow systems have been explored and used extensively in bioinformatics but there is relatively little progress in that direction in the cryo-EM community. \scipion has been designed following a strict workflow paradigm. This approach is specially useful for tasks that need to go from a given start (movies) to a defined end (3D map), but where there are many different paths (algorithms) and steps to get there.

 In a workflow approach, users (including microscope staff) can reasonably be expected to edit the paths. They can potentially play with the pipeline in ways that are not possible when having a programmer writing the code for them. Furthermore, not having the processing path hardcoded in the program means you do not have to recompile, so changes can be quickly tested   even at run time. The \scipion workflow approach provides provenance, that is,  for any created object (micrograph, CTFs, class, etc) \scipion stores metadata such as which programs, which parameters, and which exact path was followed to create any output data.  In this way,  real tracking and reproducibility are provided.

 \scipion stream capabilities can handle large and continuously produced volumes of data. The benefits of streaming are clear: (1) it helps to process data as soon as it is generated,  providing almost instant results on data quality; (2) it saves user processing time and resources at her home institutions since part of the processing is done while data is being acquired; and (3) 
 it further reduces the overall experimental time by allowing processing the data while it is transferred to the final user laboratory.

>>>>

 Workflow systems have been explored and used extensively in bioinformatics but they have raised little interest in the cryo-EM community. \scipion has been designed following a strict workflow paradigm. This approach is specially useful for tasks that need to go from a given start (movies) to a defined end (3D map), but where there are many different paths (algorithms) and steps to get there.

 Workflows implement pipelines and describe them as a set of interconnected steps. In a workflow approach, users (including microscope staff) can reasonably be expected to modify the paths and incorporate new steps using a GUI. Alternate solutions as scripts do not clearly separate each step definition from the connections between the different steps and, for no programmers, are scary and difficult to modify. The \scipion workflow approach provides provenance, that is,  for any created object (micrograph, CTFs, class, etc) \scipion stores metadata such as which programs, which parameters, and which exact path was followed to create any output data.  In this way,  real tracking and reproducibility are provided.

 \scipion stream capabilities can handle large and continuously produced volumes of data. The benefits of streaming are clear: (1) it helps to process data as soon as it is generated,  providing almost instant results on data quality; (2) it saves user processing time and resources at the user home institution since part of the processing is done while data is being acquired; and (3) 
 it further reduces the overall experimental time by allowing processing the data while it is transferred to the final user laboratory.
